### 2020-02-23 manuscript/ERL_global_C_review.Rmd
worked on intro

Despite the centrality of forest C cycling in regulating both atmospheric CO~2~ and ecosystem function and diversity, we lack a comprehensive understanding of how C cycling varies with across biomes and with stand age. While satellite measurements are increasingly useful for estimating forest $GPP$ (Bagdley et al. 2019), aboveground biomass ($B_{ag}$) (REFS), and possibly woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements. Moreover, data on stand characteristics--including species composition, age, and disturbance history--are critical to interpreting and projecting into the future, but [usually rely on ground-based measurements]. Thus, ground-based measurements are--and will continue to be--central to characterizing forest C cycling. Tens of thousands of forest C measurements have been published, but their distribution across literally thousands of scientific articles-- along with variation in data formats, units, measurement methods, *etc.*--have made them effectively inaccesible for many global-scale analyses, including those attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011), and model evaluation(Clark et al 2017, Luo et al 2012). Important progress has been made in synthesizing global data to address how C cycling varies with respect to climate (Luyssaert et al. 2007; REFS) and stand age (REFS),

*A relative dearth of data on C cycling in secondary forests, particularly in the tropics (Anderson-Teixeira et al 2016), is problematic in that almost 2/3 of the world’s forests were secondary as of 2010 (FAO 2010), implying an under-filled need to characterize age-related trends in forest C cycling.*

To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018). ForC contains data on forest ecosystem C stocks and annual fluxes (>50 variables) and associated data required for interpretation (*e.g.*, stand history, measurement methods) amalgamated from numerous previous data compilations and directly from original publications. ForC currently contains **#** records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.

@@

>>>>

Despite the centrality of forest C cycling in regulating both atmospheric CO~2~ and ecosystem function and diversity, we lack a comprehensive understanding of how C cycling varies with across biomes and with stand age. While satellite measurements are increasingly useful for estimating forest $GPP$ (Bagdley et al. 2019), aboveground biomass ($B_{ag}$) (REFS), and possibly woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements. Moreover, data on stand characteristics--including species composition, age, and disturbance history--are critical to interpreting and projecting into the future, but [usually rely on ground-based measurements]. Thus, ground-based measurements are--and will continue to be--central to characterizing forest C cycling. Tens of thousands of forest C measurements have been published, but their distribution across literally thousands of scientific articles-- along with variation in data formats, units, measurement methods, *etc.*--have made them effectively inaccesible for many global-scale analyses, including those attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011), and model evaluation (Clark et al 2017, Luo et al 2012). Important progress has been made in synthesizing data to address how C cycling varies across forest biomes (Luyssaert et al. 2007; REFS) and stand ages (REFS), yet generally consider only a limited set of C variables (but see Anderson-Teixeira et al 2016 for the tropics) and do not consider age-by-biome interactions. To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018). ForC contains data on forest ecosystem C stocks and annual fluxes (>50 variables) and associated data required for interpretation (*e.g.*, stand history, measurement methods) amalgamated from numerous previous data compilations and directly from original publications. ForC currently contains **#** records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.

@@

A relative dearth of data on C cycling in secondary forests, particularly in the tropics (Anderson-Teixeira et al 2016), is problematic in that almost 2/3 of the world’s forests were secondary as of 2010 (FAO 2010), implying an under-filled need to characterize age-related trends in forest C cycling.
